Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-13 shanghai.tex
Admissible character -> quasicharacter

\begin{abstract}
  We introduce and study the category of quasicharacter sheaves on smooth commutative
  group schemes $G$ locally of finite type over finite fields
  $\Fq$. Assuming that the geometric component group of $G$
  is finitely-generated, we show that the group of isomorphism classes
  of quasicharacter sheaves on $G$ is canonically isomorphic to the group
  of characters of $G(\Fq)$. We use this result to exhibit an
  isomorphism between admissible characters of $T(K)$ and isomorphism
  classes of quasicharacter sheaves on the Greenberg transform of the Néron
  model of $T$, where $K$ is any non-Archimdean local field and $T$ is
  any algebraic torus over $K$.
\end{abstract}

@@

The main result of this paper is a canonical isomorphism
\[
\QCiso{\GN{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of quasicharacter sheaves on
$\GN{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth less than $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \David{I'll do this when we're closer to being finished.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasi-characters of
%$K^\times$: the monoidal category of quasicharacter sheaves on the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

In this paper we show that the monoidal category of quasicharacter sheaves
on the Greenberg transform of the locally finite type Neron model of
$T$ provides an elegant, canonical model for the group of
(quasi-)characters of $T(K)$, where $K$ is any non-Archimedean local
field and $T$ is any algebraic torus over $K$.

@@

Our main result on the geometrization of admissible characters of
$p$-adic tori [insert internal reference here] follows from a general
fact established later in this paper: if the geometric component group
of $G$ is finitely-generated then the trace of Frobenius defines an
isomorphism from the Picard group of quasicharacter sheaves $G$ to the
group of characters of $G(\Fq)$ (here, as throughout the paper, $G$ is
a commutative group scheme, smooth and locally of finite type over $\Fq$).

@@

%\section{Application to admissible characters}

%\input{pro}
We may also assemble the group $\Fq$-schemes $\TT_{d}$ into one
glorious group pro-scheme over $\Fq$:
\begin{equation}\label{projlim}
  \cdef{\GN{T}} \ceq \varprojlim_{d\in \ZZ_{>0}} \GN{T}_{d}.
\end{equation}
This is the \cdef{Greenberg transform of the N\'eron model of $T$}.
 Some fundamental properties of $\GN{T}$ are:
 \begin{enumerate}
   \labitem{(GN.2)}{T(K)} $\GN{T}(\Fq) = T(K)$, canonically;
  \labitem{(GN.3)}{scheme} $\GN{T}$ is a commutative group scheme over $\Fq$.
  \end{enumerate}
This gives us a natural way to geometrize all admissible characters of $T(K)$
without reference to depth.
Let \cdef{$\QC(\GN{T})$} be the ind-category $\varinjlim_{d\in \ZZ_{>0}} \QC(\GN{T}_d)$;
we refer to objects of $\QC(\GN{T})$ as \cdef{quasicharacter sheaves on $T$}.

\begin{corollary}
Let $T$ be an algebraic torus over a non-Archimedean local field $K$.
The Picard group of the category $\QC(\GN{T})$ of quasicharacter sheaves on $T$ is canonically isomorphic to the group of admissible characters of $T(K)$:
  \[
 \QCiso{\GN{T}} \iso \Hom_\text{ad}(T(K),\EEx).
  \]
\end{corollary}

>>>>

\begin{abstract}
  We introduce and study the category of quasicharacter sheaves on smooth commutative
  group schemes $G$ locally of finite type over finite fields
  $\Fq$. Assuming that the geometric component group of $G$
  is finitely-generated, we show that the group of isomorphism classes
  of quasicharacter sheaves on $G$ is canonically isomorphic to the group
  of characters of $G(\Fq)$. We use this result to exhibit an
  isomorphism between quasicharacters of $T(K)$ and isomorphism
  classes of quasicharacter sheaves on the Greenberg transform of the Néron
  model of $T$, where $K$ is any non-Archimdean local field and $T$ is
  any algebraic torus over $K$.
\end{abstract}

@@

The main result of this paper is a canonical isomorphism
\[
\QCiso{\GN{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of quasicharacter sheaves on
$\GN{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth less than $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \David{I'll do this when we're closer to being finished.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasicharacters of
%$K^\times$: the monoidal category of quasicharacter sheaves on the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

In this paper we show that the monoidal category of quasicharacter sheaves
on the Greenberg transform of the locally finite type Neron model of
$T$ provides an elegant, canonical model for the group of
quasicharacters of $T(K)$, where $K$ is any non-Archimedean local
field and $T$ is any algebraic torus over $K$.

@@

Our main result on the geometrization of quasicharacters of
$p$-adic tori [insert internal reference here] follows from a general
fact established later in this paper: if the geometric component group
of $G$ is finitely-generated then the trace of Frobenius defines an
isomorphism from the Picard group of quasicharacter sheaves $G$ to the
group of characters of $G(\Fq)$ (here, as throughout the paper, $G$ is
a commutative group scheme, smooth and locally of finite type over $\Fq$).

@@

%\section{Application to quasicharacters}

%\input{pro}
We may also assemble the group $\Fq$-schemes $\TT_{d}$ into one
glorious group pro-scheme over $\Fq$:
\begin{equation}\label{projlim}
  \cdef{\GN{T}} \ceq \varprojlim_{d\in \ZZ_{>0}} \GN{T}_{d}.
\end{equation}
This is the \cdef{Greenberg transform of the N\'eron model of $T$}.
 Some fundamental properties of $\GN{T}$ are:
 \begin{enumerate}
   \labitem{(GN.2)}{T(K)} $\GN{T}(\Fq) = T(K)$, canonically;
  \labitem{(GN.3)}{scheme} $\GN{T}$ is a commutative group scheme over $\Fq$.
  \end{enumerate}
This gives us a natural way to geometrize all quasicharacters of $T(K)$
without reference to depth.
Let \cdef{$\QC(\GN{T})$} be the ind-category $\varinjlim_{d\in \ZZ_{>0}} \QC(\GN{T}_d)$;
we refer to objects of $\QC(\GN{T})$ as \cdef{quasicharacter sheaves on $T$}.

\begin{corollary}
Let $T$ be an algebraic torus over a non-Archimedean local field $K$.
The Picard group of the category $\QC(\GN{T})$ of quasicharacter sheaves on $T$ is canonically isomorphic to the group of quasicharacters of $T(K)$:
  \[
 \QCiso{\GN{T}} \iso \Hom_\text{ad}(T(K),\EEx).
  \]
\end{corollary}
